---
title: "Hands-on Exercise 1"
author: "Kwee Cheng"
date: "August 17, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

Edit: Update Metadata

## **1.0 Overview**

Geospatial Data Science is a process of importing, wrangling, integrating, and processing geographically referenced data sets. In this hands-on exercise, you will learn how to perform geospatial data science tasks in R by using **sf** package.

By the end of this hands-on exercise, you should acquire the following competencies:

-   installing and loading [sf](https://r-spatial.github.io/sf/) and [tidyverse](https://www.tidyverse.org/) packages into R environment,

-   importing geospatial data by using appropriate functions of **sf** package,

-   importing aspatial data by using appropriate function of **readr** package,

-   exploring the content of simple feature data frame by using appropriate **Base R** and **sf** functions,

-   assigning or transforming coordinate systems by using using appropriate *sf* functions,

-   converting an aspatial data into a sf data frame by using appropriate function of **sf** package,

-   performing geoprocessing tasks by using appropriate functions of **sf** package,

-   performing data wrangling tasks by using appropriate functions of **dplyr** package and

performing Exploratory Data Analysis (EDA) by using appropriate functions from **ggplot2** package.

## **2.0 Getting Started**

Data are key to data analytics including geospatial analytics. Hence, before analysing, we need to assemble the necessary data. In this hands-on exercise, you are required to extract the necessary data sets from the following sources:

-   Master Plan 2014 Subzone Boundary (Web) from [data.gov.sg](https://data.gov.sg/)

-   Pre-Schools Location from [data.gov.sg](https://data.gov.sg/)

-   Cycling Path from [LTADataMall](https://www.mytransport.sg/content/mytransport/home/dataMall.html)

-   Latest version of Singapore Airbnb listing data from [Inside Airbnb](http://insideairbnb.com/get-the-data.html)

### 2.1 **Extracting the geospatial data sets**

Next, at the `Hands-on_Ex01` folder, create a sub-folder called `data`. Then, inside the `data` sub-folder, create two sub-folders and name them `geospatial` and `aspatial` respectively.

Place `Master Plan 2014 Subzone Boundary (Web)`, `Pre-Schools Location` and `Cycling Path` zipped files into `geospatial` sub-folder and unzipped them. Copy the unzipped files from their respective sub-folders and place them inside `geospatial` sub-folder.

### **2.2 Extracting the aspatial data set**

Now, you will extract the downloaded `listing` data file. At `Downloads` folder, cut and paste `listing.csv` into `aspatial` sub-folder.

### 2.3 Install R-Package

In this hands-on exercise, two R packages will be used. They are:

-   **sf** for importing, managing, and processing geospatial data, and

-   **tidyverse** for performing data science tasks such as importing, wrangling and visualising data.

Tidyverse consists of a family of R packages. In this hands-on exercise, the following packages will be used:

-   **readr** for importing csv data,

-   **readxl** for importing Excel worksheet,

-   **tidyr** for manipulating data,

-   **dplyr** for transforming data, and

-   **ggplot2** for visualising data

Enter the following:

```{r}
pacman::p_load(sf, tidyverse)
```

## 3.0 Importing Geospatial Data

In this section, you will import the following geospatial data into R by using [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package:

-   `MP14_SUBZONE_WEB_PL`, a polygon feature layer in ESRI shapefile format,

-   `CyclingPath`, a line feature layer in ESRI shapefile format, and

-   `PreSchool`, a point feature layer in kml file format.

### 3.1 **Importing polygon feature data in shapefile format**

The the following code uses *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a polygon feature data frame.

```{r}
mpsz = st_read(dsn = "../Hands-on_Ex01/data/geospatial/MasterPlan", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

### **3.2 Importing polyline feature data in shapefile form**

The following code uses *st_read()* function of **sf** package to import `CyclingPath` shapefile into R as line feature data frame.

```{r}
cyclingpath = st_read(dsn = "../Hands-on_Ex01/data/geospatial/CyclingPath", 
                         layer = "CyclingPathGazette")
```

### **3.3 Importing GIS data in kml format**

The `PreSchoolsLocation` is in kml format. The following code will be used to import the kml into R.

```{r}
preschool = st_read("../Hands-on_Ex01/data/geospatial/PreSchool/PreSchoolsLocation.kml")
```

## 4.0 **Checking the Content of A Simple Feature Data Frame**

This section covers the different ways to retrieve information related to the contents of a data frame.

### 4.1 ***st_geometry()***

The column in the sf data.frame that contains the geometries is a list, of class `sfc`. We can retrieve the geometry list-column in this case by mpsz\$geom or mpsz\[\[1\]\], but the more general way uses *st_geometry()* as shown in the following:

```{r}
st_geometry(mpsz)
```

### 4.2 **glimpse()**

Beside the basic feature information, we also would like to learn more about the associated attribute information in the data frame. *glimpse()* of **dplyr**.

```{r}
glimpse(mpsz)
```

### 4.3 ***head()***

Sometimes we would like to reveal complete information of a feature object, this is the job of [*head()*](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/head) of Base R

```{r}
head(mpsz, n=5)
```

## 5.0 Plotting Geospatial Data

In geospatial data science, by looking at the feature information is not enough. We are also interested to visualise the geospatial features. This is the time you will find *plot()* of R Graphic comes in very handy as shown in the following:

```{r}
plot(mpsz)
```

The default plot of an sf object is a multi-plot of all attributes, up to a reasonable maximum as shown above. We can, however, choose to plot only the geometry by using the following:

```{r}
plot(st_geometry(mpsz))
```

Alternatively, we can also choose the plot the sf object by using a specific attribute as shown in the following:

```{r}
plot(mpsz["PLN_AREA_N"])
```

## 6.0 **Working with Projection**

Map projection is an important property of a geospatial data. In order to perform geoprocessing using two geospatial data, we need to ensure that both geospatial data are projected using similar coordinate system.

In this section, you will project a simple feature data frame form one coordinate system to another coordinate system. The technical term of this process is called ***projection transformation.***

### 6.1 **Assigning EPSG code to a simple feature data frame**

One of the common issue that can happen during importing geospatial data into R is that the coordinate system of the source data was either missing (such as due to missing .proj for ESRI shapefile) or wrongly assigned during the importing process.

An example the coordinate system of `mpsz` simple feature data frame by using *st_crs()* of *sf* package as shown in the following:

```{r}
st_crs(mpsz)
```

Although `mpsz` data frame is projected in svy21 but when we read until the end of the print, it indicates that the EPSG is 9001. This is a wrong EPSG code because the correct EPSG code for svy21 should be [3414](https://epsg.io/3414).

In order to assign the correct EPSG code to `mpsz` data frame, *st_set_crs()* of **sf** package is used as shown in the following:

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

check the CSR again by using the following:

```{r}
st_crs(mpsz3414)
```

### 6.2 **Transforming the projection of preschool from wgs84 to svy21**

In geospatial analytics, it is very common for us to transform the original data from geographic coordinate system to projected coordinate system. This is because geographic coordinate system is not appropriate if the analysis need to use distance or/and area measurements.

Let us take preschool simple feature data frame as an example. The print below reveals that it is in wgs84 coordinate system.

```         
Geometry set for 2290 features  Geometry type: POINT Dimension:     XYZ Bounding box:  xmin: 103.6878 ymin: 1.247759 xmax: 103.9897 ymax: 1.462134 z_range:       zmin: 0 zmax: 0 Geodetic CRS:  WGS 84 First 5 geometries:
```

In this scenario *st_set_crs()* is not appropriate and *st_transform()* of sf package should be used. This is because we need to reproject `preschool` from one coordinate system to another coordinate system mathemetically.

Perform the projection transformation by using the following:

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

Next, let us display the content of preschool3414 sf data frame as shown below.

```         
Geometry set for 2290 features  Geometry type: POINT Dimension:     XYZ Bounding box:  xmin: 11810.03 ymin: 25596.33 xmax: 45404.24 ymax: 49300.88 z_range:       zmin: 0 zmax: 0 Projected CRS: SVY21 / Singapore TM First 5 geometries:
```

## 7.0 **Importing and Converting An Aspatial Data**

In practice, it is not unusual that we will come across data such as `listing` of Inside Airbnb. We call this kind of data aspatial data. This is because it is not a geospatial data but among the data fields, there are two fields that capture the x- and y-coordinates of the data points.

In this section, you will learn how to import an aspatial data into R environment and save it as a tibble data frame. Next, you will convert it into a simple feature data frame.

## 7.1 **Importing the aspatial data**

Since `listings` data set is in csv file format, we will use [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) of **readr** package to import `listing.csv` as shown the code chunk below. The output R object is called `listings` and it is a [tibble data frame](https://r4ds.had.co.nz/tibbles.html).

```{r}
listings <- read_csv("../Hands-on_Ex01/data/aspatial/listings.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The following shows *list()* of Base R instead of *glimpse()*

```{r}
list(listings) 
```

### 7.2 **Creating a simple feature data frame from an aspatial data frame**

The following `listing` data frame into a simple feature data frame by using [*st_as_sf()*](https://r-spatial.github.io/sf/reference/st_as_sf.html) of **sf** packages

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

Let us examine the content of this newly created simple feature data frame.

```{r}
glimpse(listings_sf)
```

## 8.0 **Geoprocessing with sf package**

Besides providing functions to handling (i.e. importing, exporting, assigning projection, transforming projection etc) geospatial data, **sf** package also offers a wide range of geoprocessing (also known as GIS analysis) functions.

In this section, you will learn how to perform two commonly used geoprocessing functions, namely [buffering](https://www.gislounge.com/buffers-in-gis/) and point in polygon count.

### 8.1 Buffering

The scenario:

The authority is planning to upgrade the exiting cycling path. To do so, they need to acquire 5 metres of reserved land on the both sides of the current cycling path. You are tasked to determine the extend of the land need to be acquired and their total area.

The solution:

Firstly, [*st_buffer()*](https://r-spatial.github.io/sf/reference/geos_unary.html) of **sf** package is used to compute the 5-meter buffers around cycling paths

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

This is followed by calculating the area of the buffers as shown in the following:

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

Lastly, sum() of Base R will be used to derive the total land involved

```{r}
sum(buffer_cycling$AREA)
```

### 8.2 **Point-in-polygon count**

The scenario:

A pre-school service group want to find out the numbers of pre-schools in each Planning Subzone.

The solution:

The code chunk below performs two operations at one go. Firstly, identify pre-schools located inside each Planning Subzone by using [st_intersects()](https://r-spatial.github.io/sf/reference/geos_binary_pred.html). Next, [*length()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/length) of Base R is used to calculate numbers of pre-schools that fall inside each planning subzone.

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

You can check the summary statistics of the newly derived *PreSch Count* field by using *summary()* as shown in the following:

```{r}
summary(mpsz3414$`PreSch Count`)
```

To list the planning subzone with the most number of pre-school, the [*top_n()*](https://dplyr.tidyverse.org/reference/top_n.html) of **dplyr** package is used as shown in the following:

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

> #### Calculate the density of pre-school by planning subzone

Firstly, the code chunk below uses [*st_area()*](https://r-spatial.github.io/sf/reference/geos_measures.html) of **sf** package to derive the area of each planning subzone.

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

Next, [*mutate()*](https://dplyr.tidyverse.org/reference/mutate.html) of [**dplyr**](https://dplyr.tidyverse.org/) package is used to compute the density by using the following

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## 9.0 **Exploratory Data Analysis (EDA)**

In practice, many geospatial analytics start with Exploratory Data Analysis. In this section, you will learn how to use appropriate [ggplot2](https://ggplot2.tidyverse.org/) functions to create functional and yet truthful statistical graphs for EDA purposes.

Firstly, we will plot a histogram to reveal the distribution of `PreSch Density`. Conventionally, *hist()* of R Graphics will be used as shown in the following

```{r}
hist(mpsz3414$`PreSch Density`)
```

Although the syntax is very easy to use however the output is far from meeting publication quality. Furthermore, the function has limited room for further customisation.

In the code chunk below, appropriate **ggplot2** functions will be used.

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

> #### Using ggplot2 method, plot a scatterplot showing the relationship between Pre-school Density and Pre-school Count

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
