---
title: "In class exercise 2"
author: "Kwee Cheng"
date: "August 26, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## **1.0 Getting Started**

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## 2.0 Importing the data into R

[Master Plan 2014 Subzone Boundary (Web) (SHP)](https://beta.data.gov.sg/datasets/d_d14da225fccf921049ab64238ff473d9/view)\
[Master Plan 2014 Subzone Boundary (Web) (KML)](https://beta.data.gov.sg/datasets/d_5cb80a95445f236737f6bc2bfe5f159d/view)

Import Master Plan 2014 Subzone Boundary (Web) Shapefile

```{r}
mpsz <- st_read(dsn = "data/", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Write Master Plan 2014 Subzone Boundary (Web) Shapefile into KML format as the one provided is corrupted

```{r}
st_write(mpsz,"data/MP14_SUBZONE_WEB_PL.kml", delete_dsn = TRUE)
```

Import newly created KML file

```{r}
mp_kml_14 = st_read("data/MP14_SUBZONE_WEB_PL.kml")
```

Import Master Plan 2019 Subzone Boundary (Web) ShapeFile

```{r}
mpsz19 <- st_read(dsn = "data/", 
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

Import Master Plan 2019 Subzone Boundary (Web) KML

```{r}
mp_kml_19 = st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

Import Population CSV file

```{r}
popdata <- read_csv("data/respopagesextod2023.csv")
```

## 3.0 Data Wrangling

```{r}
popdata2023 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP)

colnames(popdata2023)
```

Data Processing

```{r}
popdata2023 <- popdata2023 %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[14])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+
  rowSums(.[15]))%>%
  mutate(`AGED`=rowSums(.[16:21])) %>%
  mutate(`TOTAL`=rowSums(.[3:21])) %>%  
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
  /`ECONOMY ACTIVE`) %>%
    select(`PA`, `SZ`, `YOUNG`, 
         `ECONOMY ACTIVE`, `AGED`, 
         `TOTAL`, `DEPENDENCY`)
```

Join popdata2023 and mpsz_19

```{r}
popdata2023 <- popdata2023 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper))
```

```{r}
mpsz_pop2023 <- left_join(mpsz19, popdata2023,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
pop2023_mpsz <- left_join(popdata2023, mpsz19,
                          by = c("SZ" = "SUBZONE_N"))
```
