---
title: "In Class Exercise 4"
author: "Kwee Cheng"
date: "September 9, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
categories: [In-Class, Code]
---

load packages

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse,sparr)
```

Import the study area

```{r}
kbb <- st_read(dsn="data/rawdata/",layer = "Kepulauan_Bangka_Belitung")
```

revised import code

```{r}
kbb_sf <- st_read(dsn="data/rawdata/",
                  layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

Converting to OWIN

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

check class to confirm output

```{r}
class(kbb_owin)
```

Import forest fire data

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
                        crs = 4326) %>%
  st_transform(crs = 32748)
```

convert data type of acq_date to numeric

```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label = TRUE, abbr = FALSE))
```

Overall Plot

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots()
```

visualise by months

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
tm_facets(by="Month_fac",
            free.coords=FALSE,
            drop.units = TRUE)
```

sparr able to incoorperate the control group, able to compare with the control group

Computing STKDE

extracting forest fires by month

```{r}
fire_month <- fire_sf %>%
  select(Month_num)
```

Creating ppp

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

use summary to checkout output is in the correct object class

```{r}
summary(fire_month_ppp)
```

check for duplicated

```{r}
any(duplicated(fire_month_ppp))
```

create owin object

combine origin_am_ppp and am_owin objects into one

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

Compute Spatio-temporal KDE

spattemp.density() of sparr package to compute STKDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

plotting the spatio-temporal KDE object

```{r}
#| fig-width: 12
#| fig-height: 10
tims <- c(7,8,9,10,11,12)
  par(mfcol=c(2,3))
for (i in tims){
  plot(st_kde, i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month",i))
}
```
